Approximation as Differentiation

### 2025-05-05 appendix/notes/auto-diff.tex
Correction.

Write $\tangents_x(\RR^n)$ for the tangent space at a point $x \in \RR^n$. Then the \emph{forward derivative}
(tangent map or pushforward) $\pushf{f}_x$ of a differentiable function $f: \RR^m \to \RR^n$ at $x \in \RR^m$
is the unique linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$ given by the Jacobean matrix
of $f$ at $x$. Since $\RR^n$ is flat, $\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so actually
$\pushf{f}_x$ is a linear map $\RR^m \linearto \RR^n$. The \emph{backward derivative} (cotangent map or
pullback) $\pullf{f}_x$ is the unique linear map $\RR^n \linearto \RR^m$ given by the transpose (adjoint) of
the Jacobean matrix of $f$ at $x$.

@@

For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, define $f_1 = \pi_1 \comp f$ and $f_2$ for
$\pi_2 \comp f$. Regardless of whether the linear maps $\RR^m \linearto \RR^n$ happen to be derivatives, we
can define composition on such maps as:
\begin{align*}
(g \comp f)_1(x) &= f_1(g_1(x)) \\
(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
\end{align*}
\noindent and verify that $\tangents$ is indeed a functor. If $f_2(x)$ happens to be the derivative of $f_1$
at $x$ for any $x \in \RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain
rule and $(g \comp f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.

>>>>

Write $\tangents_x(\RR^n)$ for the tangent space at a point $x \in \RR^n$. Then the \emph{forward derivative}
(tangent map or pushforward) $\pushf{f}_x$ of a differentiable function $f: \RR^m \to \RR^n$ at $x \in \RR^m$
is the unique linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$ given by the Jacobean matrix
of $f$ at $x$. Since $\RR^n$ is flat, $\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so in fact
$\pushf{f}_x$ is a linear map $\RR^m \linearto \RR^n$. The \emph{backward derivative} (cotangent map or
pullback) $\pullf{f}_x$ is the unique linear map $\RR^n \linearto \RR^m$ given by the transpose (adjoint) of
the Jacobean matrix of $f$ at $x$.

@@

For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, write $f_1$ for $\pi_1 \comp f$ and $f_2$ for
$\pi_2 \comp f$. Regardless of whether the linear maps $\RR^m \linearto \RR^n$ happen to be derivatives, we
can define composition on such maps as:
\begin{align*}
(g \comp f)_1(x) &= f_1(g_1(x)) \\
(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
\end{align*}
\noindent and verify that $\tangents$ is indeed a functor. If $f_2(x)$ happens to be the derivative of $f_1$
at $x$ for any $x \in \RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain
rule and $(g \comp f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.
